---
title: "STAT697MVFinal"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = F}
library(tidyverse)
library(DataExplorer)
library(zoo)
library(plotly)
library(lubridate)
library(scales)
library(lubridate)
library(corrplot)
library(psych)
library(ggplot2)
library(pls)
library(zoo)
library(Metrics)
library(VGAM)
library(e1071)
library(mvtnorm)
library(sn)
data <- read_csv("AirQualityUCI.csv")
```

```{r}
# remove columns that don't have any title
data <- data[1:15]

# here we see that the last 114 rows are completely NAs
#data[is.na(data$Date) ==TRUE, ] 
# remove lines full of NAs
data <- data[is.na(data$Date) == FALSE, ]
```


### Context

Source: https://archive.ics.uci.edu/ml/datasets/Air+quality#

The dataset contains 9358 instances of hourly averaged responses from an array of 5 metal oxide chemical sensors embedded in an Air Quality Chemical Multi-sensor device. There are 13 attributes which potentially can be used as predictors with two outcome variables - Relative Humidity and Absolute Humidity.


1) Date (DD/MM/YYYY)

2) Time (HH.MM.SS)

3) CO(GT): True hourly averaged concentration CO in mg/m^3 (reference analyzer)

4) PT08.S1(CO): PT08.S1 (tin oxide) hourly averaged sensor response (nominally CO targeted)

5) NMHC(GT): True hourly averaged overall Non Metanic HydroCarbons concentration in microg/m^3 (reference analyzer)

6) C6H6(GT): True hourly averaged Benzene concentration in microg/m^3 (reference analyzer)

7) PT08.S2(NMHC): PT08.S2 (titania) hourly averaged sensor response (nominally NMHC targeted)

8) NOx(GT): True hourly averaged NOx concentration in ppb (reference analyzer)

9) PT08.S3(NOx): PT08.S3 (tungsten oxide) hourly averaged sensor response (nominally NOx targeted)

10) NO2(GT): True hourly averaged NO2 concentration in microg/m^3 (reference analyzer)

11) PT08.S4(NO2): PT08.S4 (tungsten oxide) hourly averaged sensor response (nominally NO2 targeted)

12) PT08.S5(O3): PT08.S5 (indium oxide) hourly averaged sensor response (nominally O3 targeted)

13) T: Temperature in °C

14) RH: Relative Humidity (%)

15) AH: Absolute Humidity


## 1) Introduce the data
```{r}
# 9357 rows
# 13 continuous variables + column Date and column Time
# no NAs
t(introduce(data))
str(data)
```

## 2) Missing values
```{r}
#Missing values are tagged with -200 value. 
data[data == -200] <- NA

# NMHC(GT) has 8443 missing values so drop it
colSums(is.na(data))
data <- subset(data, select = -c(`NMHC(GT)`))
# delete 366 rows with mostly NAs
data <- data[-as.vector(which(is.na(data$`PT08.S3(NOx)`))),]

### Replace NAs with the mean value of the column
data[3:14] <- na.aggregate(data[3:14], by=1)
```

## 3) Transform Date type
```{r}
data$Date <- mdy(data$Date)
data$Time <- hms(data$Time)
```

## 4) Description of the data
```{r, message=F}
summary(data)
describe(data)
```


## 5) Summary statistics
```{r}
# Histogram and density plots to analyze/represent continuous variables 
plot_density(data)
plot_histogram(data)
```

## 6) Correlations
```{r}
data_for_cor <- data %>%
  mutate(hour= hour(Time), month = month(Date))
data_for_cor <- data_for_cor[3:16]

# the "pairwise.complete.obs" makes sure to avoid NAs
corrs = cor(data_for_cor, use="pairwise.complete.obs")
corrplot(corrs, type="upper", method="color", addCoef.col = "black")
```

## 7) RH and AH over time
```{r}
## Over Date
plot_ly(data, x = ~as.Date(Date)) %>%
  add_lines(y = ~RH, 
            name = 'RH', 
            mode = 'lines') %>%
  add_lines(y = ~AH, 
            name = 'AH', 
            mode = 'lines') %>%
  layout(title='RH and AH for each month of the year',
         xaxis = list(title = 'Date'),
         yaxis = list(title = 'RH and AH'))

## Over Date with scaled values
data %>%
  mutate(scl_RH = (RH - min(RH))/(max(RH) - min(RH)), scl_AH = ((AH - min(AH))/(max(AH) - min(AH))))%>%
plot_ly(x = ~as.Date(Date)) %>%
  add_lines(y = ~scl_RH, 
            name = 'RH', 
            mode = 'lines') %>%
  add_lines(y = ~scl_AH, 
            name = 'AH', 
            mode = 'lines') %>%
  layout(title='Scaled RH and AH for each month of the year',
         xaxis = list(title = 'Date'),
         yaxis = list(title = 'scaled RH and AH'))
```


```{r}
#### PER MONTH ####
# boxplot of RH for each month of the year
data %>%
 #mutate(month = month(as.POSIXlt(Date, format="%d/%m/%Y"))) %>%
  mutate(month= as.factor(month(Date))) %>%
  mutate(months= as.factor(months(Date))) %>%
  group_by(month, months, RH) %>%
  ungroup() %>%
  plot_ly(x = ~month, y = ~RH, name = ~months, type = "box")%>%
  layout(title='Boxplot of RH for each month of the year ')

# boxplot of AH for each month of the year
data %>%
 # mutate(month = month(as.POSIXlt(Date, format="%d/%m/%Y"))) %>%
  mutate(month= as.factor(month(Date))) %>%
  mutate(months= as.factor(months(Date))) %>%
  group_by(month, months, AH) %>%
  ungroup() %>%
  plot_ly(x = ~month, y = ~AH, name = ~months, type = "box")%>%
  layout(title='Boxplot of AH for each month of the year ')
```


```{r}
#### PER MONTH ####
# Histogram of RH for each hour of the day
data %>%
  mutate(month= as.factor(month(Date))) %>%
  group_by(month)%>%
  summarize(mean_RH = mean(RH, na.rm = T)) %>% 
  plot_ly(x = ~month, 
          y = ~mean_RH)%>%
  layout(title='Histogram of RH for each hour of the day',
         xaxis = list(title = 'hour'),
         yaxis = list(title = 'RH'))
# Histogram of AH for each hour of the day
data %>%
  mutate(month= as.factor(month(Date))) %>%
  group_by(month)%>%
  summarize(mean_AH = mean(AH, na.rm = T)) %>% 
  plot_ly(x = ~month, 
          y = ~mean_AH)%>%
  layout(title='Histogram of AH for each hour of the day',
         xaxis = list(title = 'hour'),
         yaxis = list(title = 'AH'))
```


```{r}
#### PER MONTH ####
# Histogram RH and AH mean values per hour
data %>%
  mutate(month= as.factor(month(Date))) %>%
  group_by(month)%>%
  summarize(mean_RH = mean(RH, na.rm = T), mean_AH = mean(AH, na.rm = T)) %>%
  plot_ly(x = ~month, 
        y = ~mean_RH, 
        type = 'bar', 
        name = 'RH',
        marker = list(color = 'rgb(55, 83, 109)')) %>% 
  add_trace(y = ~mean_AH, name = 'AH', marker = list(color = 'rgb(26, 118, 255)')) %>% 
  layout(title = 'Mean values of RH and AH per month',
         xaxis = list(
           title = "Hour",
           tickfont = list(
             size = 14,
             color = 'rgb(107, 107, 107)')),
         yaxis = list(
           title = 'Mean values of RH and AH',
           titlefont = list(
             size = 16,
             color = 'rgb(107, 107, 107)'),
           tickfont = list(
             size = 14,
             color = 'rgb(107, 107, 107)')),
         legend = list(bgcolor = 'rgba(255, 255, 255, 0)', bordercolor = 'rgba(255, 255, 255, 0)'),
         barmode = 'group', bargap = 0.15, bargroupgap = 0.1)

# Histogram RH and AH scaled mean values per month
data %>%
 mutate(month= as.factor(month(Date))) %>%
  group_by(month)%>%
  summarize(mean_RH = mean(RH, na.rm = T), mean_AH = mean(AH, na.rm = T)) %>%
  summarize(month, scl_mean_RH = (mean_RH - min(mean_RH))/(max(mean_RH) - min(mean_RH)), scl_mean_AH = ((mean_AH - min(mean_AH))/(max(mean_AH) - min(mean_AH))))%>%
  plot_ly(x = ~month, 
        y = ~scl_mean_RH, 
        type = 'bar', 
        name = 'RH',
        marker = list(color = 'rgb(55, 83, 109)')) %>% 
  add_trace(y = ~scl_mean_AH, name = 'AH', marker = list(color = 'rgb(26, 118, 255)')) %>% 
  layout(title = 'Mean values of RH and AH per hour',
         xaxis = list(
           title = "Hour",
           tickfont = list(
             size = 14,
             color = 'rgb(107, 107, 107)')),
         yaxis = list(
           title = 'Mean values of RH and AH',
           titlefont = list(
             size = 16,
             color = 'rgb(107, 107, 107)'),
           tickfont = list(
             size = 14,
             color = 'rgb(107, 107, 107)')),
         legend = list(bgcolor = 'rgba(255, 255, 255, 0)', bordercolor = 'rgba(255, 255, 255, 0)'),
         barmode = 'group', bargap = 0.15, bargroupgap = 0.1)
```

```{r}
#### PER HOUR ####
# boxplot of RH for each hour of the day
data %>% 
  mutate(hour= as.factor(hour(Time))) %>%
  group_by(hour, RH) %>%
  ungroup() %>%
  plot_ly(x = ~hour, y = ~RH, type = "box")%>%
  layout(title='Boxplot of RH for each hour of the day')

# boxplot of AH for each hour of the day
data %>% 
  mutate(hour= as.factor(hour(Time))) %>%
  group_by(hour, AH) %>%
  ungroup() %>%
  plot_ly(x = ~hour, y = ~AH, type = "box")%>%
  layout(title='Boxplot of AH for each hour of the day')
```


```{r}
#### PER HOUR ####
# Histogram of RH for each hour of the day
data %>%
  mutate(hour= as.factor(hour(Time))) %>%
  group_by(hour) %>%
  summarize(mean_RH = mean(RH, na.rm = T)) %>% 
  plot_ly(x = ~hour, 
          y = ~mean_RH)%>%
  layout(title='Histogram of RH for each hour of the day',
         xaxis = list(title = 'hour'),
         yaxis = list(title = 'RH'))
# Histogram of AH for each hour of the day
data %>%
  mutate(hour= as.factor(hour(Time))) %>%
  group_by(hour) %>%
  summarize(mean_AH = mean(AH, na.rm = T)) %>% 
  plot_ly(x = ~hour, 
          y = ~mean_AH)%>%
  layout(title='Histogram of AH for each hour of the day',
         xaxis = list(title = 'hour'),
         yaxis = list(title = 'AH'))
```


```{r}
#### PER HOUR ####
# Histogram RH and AH mean values per hour
data %>%
  mutate(hour= as.factor(hour(Time))) %>%
  group_by(hour) %>%
  summarize(mean_RH = mean(RH, na.rm = T), mean_AH = mean(AH, na.rm = T)) %>%
  plot_ly(x = ~hour, 
        y = ~mean_RH, 
        type = 'bar', 
        name = 'RH',
        marker = list(color = 'rgb(55, 83, 109)')) %>% 
  add_trace(y = ~mean_AH, name = 'AH', marker = list(color = 'rgb(26, 118, 255)')) %>% 
  layout(title = 'Mean values of RH and AH per hour',
         xaxis = list(
           title = "Hour",
           tickfont = list(
             size = 14,
             color = 'rgb(107, 107, 107)')),
         yaxis = list(
           title = 'Mean values of RH and AH',
           titlefont = list(
             size = 16,
             color = 'rgb(107, 107, 107)'),
           tickfont = list(
             size = 14,
             color = 'rgb(107, 107, 107)')),
         legend = list(bgcolor = 'rgba(255, 255, 255, 0)', bordercolor = 'rgba(255, 255, 255, 0)'),
         barmode = 'group', bargap = 0.15, bargroupgap = 0.1)

# Histogram RH and AH scaled mean values per hour
data %>%
  mutate(hour= as.factor(hour(Time))) %>%
  group_by(hour) %>%
  summarize(mean_RH = mean(RH), mean_AH = mean(AH))%>%
  summarize(hour, scl_mean_RH = (mean_RH - min(mean_RH))/(max(mean_RH) - min(mean_RH)), scl_mean_AH = ((mean_AH - min(mean_AH))/(max(mean_AH) - min(mean_AH))))%>%
  plot_ly(x = ~hour, 
        y = ~scl_mean_RH, 
        type = 'bar', 
        name = 'RH',
        marker = list(color = 'rgb(55, 83, 109)')) %>% 
  add_trace(y = ~scl_mean_AH, name = 'AH', marker = list(color = 'rgb(26, 118, 255)')) %>% 
  layout(title = 'Mean values of RH and AH per hour',
         xaxis = list(
           title = "Hour",
           tickfont = list(
             size = 14,
             color = 'rgb(107, 107, 107)')),
         yaxis = list(
           title = 'Mean values of RH and AH',
           titlefont = list(
             size = 16,
             color = 'rgb(107, 107, 107)'),
           tickfont = list(
             size = 14,
             color = 'rgb(107, 107, 107)')),
         legend = list(bgcolor = 'rgba(255, 255, 255, 0)', bordercolor = 'rgba(255, 255, 255, 0)'),
         barmode = 'group', bargap = 0.15, bargroupgap = 0.1)
```


## 8) Multivariate Regression
```{r}
# Fitting a multiple outcomes regression model
data_ml = data[, -c(1:2)]

# We split our data into 70% train and 30% test
n <- nrow(data_ml)
cutoff <- round(.70 * n) - 1
train_data = data_ml[c(1:cutoff), ]
test_data = data_ml[c((cutoff+1):n), ]
test_x = test_data[, -c(11:12)]
test_y = test_data[, c(11:12)]
```


```{r}
# Fitting a multivariate Regression Model to our data
lm_model1 = lm(cbind(RH, AH) ~ ., data = train_data)
summary(lm_model1)

# Predicting on our test set and calculating RMSE of our predictions
pred_y = predict(lm_model1, test_x)
pred_y = as.data.frame(pred_y)
# RMSE for the outcome RH
(rmse_lm_RH = sqrt(sum((test_y[, 1] - pred_y[, 1])^2)/nrow(test_y)))
# RMSE for the outcome AH
(rmse_lm_AH = sqrt(sum((test_y[, 2] - pred_y[, 2])^2)/nrow(test_y)))

res_lm_RH = lm_model1$residuals[, 1]
res_lm_AH = lm_model1$residuals[, 2]

# Plot of residuals for outcome RH
plot(res_lm_RH)
# Plot of residuals for outcome AH
plot(res_lm_AH)

par(mfrow=c(2,1))
# Histogram of residuals for outcome RH
hist(res_lm_RH, prob=TRUE, main = "Histogram of residuals for RH",
     xlab = "Residuals - RH")
lines(density(res_lm_RH), col="red", lwd=2) # add a density estimate with defaults
lines(density(res_lm_RH, adjust=2), lty="dotted", col="darkgreen", lwd=2) 
# Histogram of residuals for outcome AH
hist(res_lm_AH, prob=TRUE, main = "Histogram of residuals for AH",
     xlab = "Residuals - AH")
lines(density(res_lm_AH), col="red", lwd=2) # add a density estimate with defaults
lines(density(res_lm_AH, adjust=2), lty="dotted", col="darkgreen", lwd=2) 
```


## 9) PCA Regression
```{r}
# Fitting a PCA on the explanatory variables
pca1 = princomp(train_data[, c(1:10)], cor = TRUE)
summary(pca1)
screeplot(pca1, type = "lines")

# Fitting a principal component regression model on the whole dataset
pcr_model = pcr(cbind(RH, AH) ~., data = train_data, scale = TRUE, validation = "CV")
summary(pcr_model)
validationplot(pcr_model)

# Basis the scree plot we select ncomp = 3 and then fit our model on the training data
pcr_model_3 = pcr(cbind(RH, AH) ~., data = train_data, scale = TRUE, validation = "CV",
                 ncomp = 3)
summary(pcr_model_3)
#validationplot(pcr_model_3)

# Predicting on our test set and calculating RMSE of our predictions for PCR model with ncomp = 3
pcr_pred = as.data.frame(predict(pcr_model_3, test_x, ncomp = 3))
# RMSE for the outcome RH
(rmse_PCA_RH = sqrt(sum((test_y[, 1] - pcr_pred[, 1])^2)/nrow(test_y)))
# RMSE for the outcome AH
(rmse_PCA_AH = sqrt(sum((test_y[, 2] - pcr_pred[, 2])^2)/nrow(test_y)))
```


## 10) Stimulation study

```{r}
## Collect the estimated betas from the MVN
r <- length(lm_model1$coefficients)/2
for (i in 1:r-1){
  assign(paste0("beta_", i), as.vector(lm_model1$coefficients[i+1,]))
}
beta <- cbind(beta_0, beta_1, beta_2, beta_3, beta_4, beta_5, beta_6, beta_7, beta_8, beta_9, beta_10)
colnames(beta) <- NULL

## Parameters for the Skew Normal Errors with choice of sigma
mu_0 <- matrix(c(0, 0), ncol = 1)
Sigma <- matrix(c(var(res_lm_RH), cor(res_lm_RH, res_lm_AH),
                  cor(res_lm_RH, res_lm_AH), var(res_lm_AH)), nrow = 2)
cpM <- list(mean=c(mu_0), var.cov=Sigma, gamma1=c(skewness(res_lm_RH)/2, skewness(res_lm_AH)/2))
dpM <- cp2dp(cpM, family="SN")

## Stimulation
niter = 1000 # number of simulation iteration
betas_MVN_RH = betas_MVN_AH = betas_MSN_RH = betas_MSN_AH = matrix(0, niter, 11)
resi_MVN_RH = resi_MVN_AH = resi_MSN_RH = resi_MSN_AH= array(0, c(n, niter))

set.seed(10) # seed number is fixed as 10, to make simulation study reproducible

X <- as.matrix(cbind(replicate(n, 1),as.matrix(data_ml[,1:10]))) # use data from the dataset
names <- colnames(X)  # store the names for the graphs
names[1] <- "Intercept"
colnames(X) <- NULL

for (iter in 1:niter){
  
  # Program for data generating based on the simulation setup
  errorMVN = mvtnorm::rmvnorm(n, c(mu_0), Sigma) 
  errorMSN = sn::rmsn(n, dp=dpM)
  
  Y_MVN = X %*% t(beta) + errorMVN
  Y_MSN = X %*% t(beta) + errorMSN
  
  # Fit regression model with generated data
  reg_MVN = lm(Y_MVN ~ X -1)
  reg_MSN = lm(Y_MSN ~ X -1)
  
  # Collect estimates such as betas and residuals
  ### MVN ###
  betas_MVN_RH[iter,] = reg_MVN$coefficients[,1]
  betas_MVN_AH[iter,] = reg_MVN$coefficients[,2]
  resi_MVN_RH[,iter] = reg_MVN$residuals[,1]
  resi_MVN_AH[,iter] = reg_MVN$residuals[,2]
  ### MSN ###
  betas_MSN_RH[iter,] = reg_MSN$coefficients[,1]
  betas_MSN_AH[iter,] = reg_MSN$coefficients[,2]
  resi_MSN_RH[,iter] = reg_MSN$residuals[,1]
  resi_MSN_AH[,iter] = reg_MSN$residuals[,2]
}


## Table for parameter estimates from Normal and Skew-Normal Errors for RH and AH
table_RH <- tibble('Estimated value for RH' =names,
                   'True Value' = rep(NA_real_, 11),
                   "Normal Estimate" = NA_real_,
                   'MVN Squared error' = NA_real_,
                   "Skew Normal Estimate" = NA_real_,
                   'MSN Squared error' = NA_real_)
for (i in 1:11){
  table_RH[i,2] <- beta[1,i]
  table_RH[i,3] <- mean(betas_MVN_RH[,i])
  table_RH[i,5] <- mean(betas_MSN_RH[,i])
}
table_RH$`MVN Squared error` <- (table_RH$`True Value` - table_RH$`Normal Estimate`)^2
table_RH$`MSN Squared error` <- (table_RH$`True Value` - table_RH$`Skew Normal Estimate`)^2
print(table_RH)

table_AH <- tibble('Estimated value for AH' =names,
                   'True Value' = rep(NA_real_, 11),
                   "Normal Estimate" = NA_real_,
                   'MVN Squared error' = NA_real_,
                   "Skew Normal Estimate" = NA_real_,
                   'MSN Squared error' = NA_real_)

for (i in 1:11){
  table_AH[i,2] <- beta[2,i]
  table_AH[i,3] <- mean(betas_MVN_AH[,i])
  table_AH[i,5] <- mean(betas_MSN_AH[,i])
}

table_AH$`MVN Squared error` <- (table_AH$`True Value` - table_AH$`Normal Estimate`)^2
table_AH$`MSN Squared error` <- (table_AH$`True Value` - table_AH$`Skew Normal Estimate`)^2
print(table_AH)



## Sampling distribution of the $\beta$s for RH and AH: Normal Errors (green) and Skew Normal Errors (blue) 
par(mfrow=c(4,3))
for (i in 1:11){
  plot(density(betas_MVN_RH[, i]), main = paste0('Density plot of beta', i-1,' (', names[i],') ', 'for RH'), col="green") + lines(density(betas_MSN_RH[, i]), col="blue", lwd=2) + abline(v = beta[1,i])
}

par(mfrow=c(4,3))
for (i in 1:11){
  plot(density(betas_MVN_AH[, i]), main = paste0('Density plot of beta', i-1,' (', names[i],') ', 'for RH'), col="green") + lines(density(betas_MSN_AH[, i]), col="blue", lwd=2) + abline(v = beta[2,i])
}


## Residuals for outcomes RH and AH with Normal and Skew Normal Errors
par(mfrow=c(2,2))
hist(resi_MVN_RH, prob=TRUE, main = "Residuals of RH with Normal Errors",
     xlab = "Residuals")
hist(resi_MVN_AH, prob=TRUE, main = "Residuals of AH with Normal Errors",
     xlab = "Residuals")
hist(resi_MSN_RH, prob=TRUE, main = "Residuals of RH with Skew-Normal Errors",
     xlab = "Residuals")
hist(resi_MSN_AH, prob=TRUE, main = "Residuals of AH with Skew-Normal Errors",
     xlab = "Residuals")
```



## 11) Principal Factor Analysis 
```{r}
data_fa <- data[3:12]
```


```{r}
### Correlation matrix ###
R <- cor(data_fa)
as.table(R)
```


### Principal component method With correlation matrix
```{r}
##### Principal Component method (with correlation matrix) ##### 

### Proportion of variation - Choice of m ###
eig <- eigen(R)
cumsum(eig$values / sum(eig$values))
```

```{r}
screeplot(princomp(data_fa), type='lines')
```


```{r}
### PC with and without rotation ###
(factors_data <- principal(r = R, nfactors = 3, rotate = "none", covar = FALSE))
(factors_data_varimax <- principal(r = R, nfactors = 2, rotate = "Varimax", covar = FALSE))
```


#### Maximum Likelihood method with correlation matrix

```{r}
##### Maximum Likelihood method (with correlation matrix) ##### 

### Proportion of variation - Choice of m ###
output.df = data.frame()
for (i in (2:ncol(data_fa))){
  ML_factors_data <- fa(r = R, fm = "ml", nfactors = i, rotate = "none", covar = FALSE)
  ML_factors_data_Varimax <- fa(r = R, fm = "ml", nfactors = i, rotate = "Varimax", covar = FALSE)
  output.df <- rbind(output.df, data.frame(m = i, 'cumulative proportion' = ML_factors_data$Vaccounted[5,i], 'cumulative proportion with rotation' = ML_factors_data_Varimax$Vaccounted[5,i]))
}
output.df
```


```{r}
### ML with and without rotation ### 
(factor_scores <- fa(r = R, fm = "ml", nfactors = 3, rotate = "none", covar = FALSE))
(factor_scores_varimax <- fa(r = R, fm = "ml", nfactors = 3, rotate = "Varimax", covar = FALSE))
```




## 12) Change point analysis 

1) changepoint
2) Segmented
3) Tree
4) bcp

```{r}
ts_data <- data
ts_data <- data %>%
  mutate(Month =  month(ts_data$Date), Hour = hour(ts_data$Time))
ts_data <- subset(ts_data, select = c(Date, Month, Hour, RH, AH))
ts_data
```


1) changepoint

```{r}
library(changepoint)
###### MONTH ###### 
## Changepoint for RH per month
X <- data.frame(date = ts_data$Date, RH = ts_data$RH)
a <- changepoint::cpt.mean(X$RH)
m <- unlist(a@param.est)
cpt <- a@cpts[1]
(X$date[cpt])
summary(a)
ggplot() +
  geom_line(data = X, aes(x = date, y = RH)) +
  geom_line(
    data = data.frame(d = c(X$date[1], X$date[cpt]), y = rep(m[1], 2)),
    aes(x = d, y = y),
    color = "red"
    ) +
  geom_line(
    data = data.frame(d = c(X$date[cpt], X$date[dim(X)[1]]), y = rep(m[2], 2)),
    aes(x = d, y = y),
    color = "red"
  )

## Changepoint for AH per month
X <- data.frame(date = ts_data$Date, AH = ts_data$AH)
a <- changepoint::cpt.mean(X$AH)
m <- unlist(a@param.est)
cpt <- a@cpts[1]
(X$date[cpt])

ggplot() +
  geom_line(data = X, aes(x = date, y = AH)) +
  geom_line(
    data = data.frame(d = c(X$date[1], X$date[cpt]), y = rep(m[1], 2)),
    aes(x = d, y = y),
    color = "red"
    ) +
  geom_line(
    data = data.frame(d = c(X$date[cpt], X$date[dim(X)[1]]), y = rep(m[2], 2)),
    aes(x = d, y = y),
    color = "red"
  )

###### HOUR ###### 
## RH
# Mean RH per hour


X = aggregate(ts_data$RH, by=list(ts_data$Hour), FUN=mean)
X <- data.frame(date = X[,1], RH = X[,2])
a <- changepoint::cpt.mean(X$RH)
m <- unlist(a@param.est)
cpt <- a@cpts[1]
(X$date[cpt])
summary(a)
ggplot() +
  geom_line(data = X, aes(x = date, y = RH)) +
  geom_line(
    data = data.frame(d = c(X$date[1], X$date[cpt]), y = rep(m[1], 2)),
    aes(x = d, y = y),
    color = "red"
    ) +
  geom_line(
    data = data.frame(d = c(X$date[cpt], X$date[dim(X)[1]]), y = rep(m[2], 2)),
    aes(x = d, y = y),
    color = "red"
  )


## AH
# Mean AH per hour
X = aggregate(ts_data$AH, by=list(ts_data$Hour), FUN=mean)
X <- data.frame(date = X[,1], AH = X[,2])
a <- changepoint::cpt.mean(X$AH)
m <- unlist(a@param.est)
cpt <- a@cpts[1]
(X$date[cpt])
summary(a)
ggplot() +
  geom_line(data = X, aes(x = date, y = AH)) +
  geom_line(
    data = data.frame(d = c(X$date[1], X$date[cpt]), y = rep(m[1], 2)),
    aes(x = d, y = y),
    color = "red"
    ) +
  geom_line(
    data = data.frame(d = c(X$date[cpt], X$date[dim(X)[1]]), y = rep(m[2], 2)),
    aes(x = d, y = y),
    color = "red"
  )

```


2) segmented

```{r}
library(segmented)
###### MONTH ###### 
## RH
fit_lm = lm(RH ~ 1, data = ts_data)  # intercept-only model
fit_segmented = segmented(fit_lm, seg.Z = ~Month, npsi = 1)  # Two change points along x
summary(fit_segmented)
plot(fit_segmented)
points(ts_data$AH)
lines.segmented(fit_segmented)
points.segmented(fit_segmented)

## AH
fit_lm = lm(AH ~ 1, data = ts_data)  # intercept-only model
fit_segmented = segmented(fit_lm, seg.Z = ~Month, npsi = 2)  # Two change points along x
summary(fit_segmented)
plot(fit_segmented)
#plot(ts_data$RH)
lines.segmented(fit_segmented, col = 'red')
points.segmented(fit_segmented, col = 'red')

###### Hour ###### 

## RH
fit_lm = lm(RH ~ 1, data = ts_data)  # intercept-only model
fit_segmented = segmented(fit_lm, seg.Z = ~Hour, npsi = 2)  # Two change points along x
summary(fit_segmented)
plot(fit_segmented)
points(ts_data$AH)
lines.segmented(fit_segmented, col = 'red')
points.segmented(fit_segmented, col = 'red')

## AH
fit_lm = lm(AH ~ 1, data = ts_data)  # intercept-only model
fit_segmented = segmented(fit_lm, seg.Z = ~Hour, npsi = 2)  # Two change points along x
summary(fit_segmented)
plot(fit_segmented)
points(ts_data$AH)
lines.segmented(fit_segmented, col = 'red')
points.segmented(fit_segmented, col = 'red')
```


3) Tree

```{r}
library(tree)
###### MONTH ###### 
## RH
tree <- tree(RH ~ Month, data = ts_data)
summary(tree)
plot(tree)
text(tree, pretty = 0)

## AH
tree <- tree(AH ~ Month, data = ts_data)
summary(tree)
plot(tree)
text(tree, pretty = 0)


###### Hour ###### 
## RH
tree <- tree(RH ~ Hour, data = ts_data)
summary(tree)
plot(tree)
text(tree, pretty = 0)

# ## AH
# tree <- tree(AH ~ Hour, data = ts_data)
# summary(tree)
# plot(tree)
# text(tree, pretty = 0)
```

4) bcp

```{r}
library(bcp)
#### MONTH ####

# Mean RH per hour
tem = aggregate(ts_data$RH, by=list(ts_data$Month), FUN=mean)
Mean_RH = tem[,2]
names(Mean_RH) = tem[,1]
#barplot(Mean_RH, las=2)
series_RH = ts(Mean_RH,start=c(0), end = c(12))


#### bcp
# Also this method identified the correct month of change.
# The lower posterior probability plot shows that at many locations, the probability of a change is very high. 
bcp_x <- bcp(series_RH, return.mcmc = TRUE)
plot(bcp_x)
# We can get the exact locations where probabilities are high (e.g. > 70%) with this code:
bcp_sum <- as.data.frame(summary(bcp_x))
# Let's filter the data frame and identify the month:
bcp_sum$id <- 1:length(series_RH)
(sel <- bcp_sum[which(bcp_x$posterior.prob > 0.7), ])
# Get the year:
time(series_RH)[sel$id] 


# Mean AH per hour
tem = aggregate(ts_data$AH, by=list(ts_data$Month), FUN=mean)
Mean_AH = tem[,2]
names(Mean_AH) = tem[,1]
#barplot(Mean_AH, las=2)
series_AH = ts(Mean_AH,start=c(0), end = c(12))


#### bcp
# Also this method identified the correct month of change.
# The lower posterior probability plot shows that at many locations, the probability of a change is very high. 
bcp_x <- bcp(series_AH, return.mcmc = TRUE)
plot(bcp_x)
# We can get the exact locations where probabilities are high (e.g. > 70%) with this code:
bcp_sum <- as.data.frame(summary(bcp_x))
# Let's filter the data frame and identify the month:
bcp_sum$id <- 1:length(series_AH)
(sel <- bcp_sum[which(bcp_x$posterior.prob > 0.7), ])
# Get the year:
time(series_RH)[sel$id] 


##### HOUR #####

# Mean RH per hour
tem = aggregate(ts_data$RH, by=list(ts_data$Hour), FUN=mean)
Mean_RH = tem[,2]
names(Mean_RH) = tem[,1]
#barplot(Mean_RH, las=2)
series_RH = ts(Mean_RH,start=c(0), end = c(23))


#### bcp
# Also this method identified the correct month of change.
# The lower posterior probability plot shows that at many locations, the probability of a change is very high. 
bcp_x <- bcp(series_RH, return.mcmc = TRUE)
plot(bcp_x)
# We can get the exact locations where probabilities are high (e.g. > 70%) with this code:
bcp_sum <- as.data.frame(summary(bcp_x))
# Let's filter the data frame and identify the month:
bcp_sum$id <- 1:length(series_RH)
(sel <- bcp_sum[which(bcp_x$posterior.prob > 0.7), ])
# Get the year:
time(series_RH)[sel$id] 


# Mean AH per hour
tem = aggregate(ts_data$AH, by=list(ts_data$Hour), FUN=mean)
Mean_AH = tem[,2]
names(Mean_AH) = tem[,1]
#barplot(Mean_AH, las=2)
series_AH = ts(Mean_AH,start=c(0), end = c(23))


#### bcp
# Also this method identified the correct month of change.
# The lower posterior probability plot shows that at many locations, the probability of a change is very high. 
bcp_x <- bcp(series_AH, return.mcmc = TRUE)
plot(bcp_x)
# We can get the exact locations where probabilities are high (e.g. > 70%) with this code:
bcp_sum <- as.data.frame(summary(bcp_x))
# Let's filter the data frame and identify the month:
bcp_sum$id <- 1:length(series_AH)
(sel <- bcp_sum[which(bcp_x$posterior.prob > 0.7), ])
# Get the year:
time(series_RH)[sel$id] 
```


